---
title: "Eng Aux 2"
author: "Phillip.Escandon"
date: "Nov 27, 2017"
output: 
  html_notebook: 
    toc: yes
---
# Eng Aux Data files
###**eng_aux_podins.dat**  
Column / Field  
1:      timestamp  
2:      valid 0=data not valid,1 = valid  
7:      eastVel  
8:      northVel  
9:      downVel  

###**eng_aux_podstateA.dat**  
Column / Field  
1:     timestamp  
8:     mach speed  
94:    gpstime    

###**eng_aux_podstateD.dat**  
Column / Field  
1:  timestamp  
15: FOV Stat: 0xaa = NFOV, 0x6a=NFOV 0x55 WFOV 0x95 SWFOV  
16: shutter status

    0 = open  
    1 =  closed  
    2 = opening  
    3 = closing  
    4 = stuck closed  
    5 = stuck open  
    6 = in transit  
    7 = out of position  
    8 = error - open and closed  
    9 = unknown  
    10 = stuck in transit  
21. sensorID  
  

  
###**eng_aux_recce_task.dat**  
Column / Field  
1 : timestamp  
8 : mpUserID task name known by user  
272 : groundSpeed Ground speed at start gate m/s  
  
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Engineering Auxillary Data

```{r Load Library}
library(tidyverse)
library(plotly)
```

#### Gather Data
```{r Read in data, include=FALSE}

podINS <- read_table2('../data/15090103s2/SSD_SEQ_1/eng_aux_podins.dat',col_names = FALSE)

podStateA <- read_table2('../data/15090103s2/SSD_SEQ_1/eng_aux_podstatea.dat',col_names = FALSE)

podStateD <- read_table2('../data/15090103s2/SSD_SEQ_1/eng_aux_podstated.dat',col_names = FALSE)

recceTask <- read_table2('../data/15090103s2/SSD_SEQ_1/eng_aux_recce_task.dat',col_names = FALSE)
```

#### Select only the appropriate columns indicated above  

```{r Select Needed Columns}
podINSFiltered <- select(podINS,1,2,7,8,9)
podStateAFiltered <- select(podStateA,1,8,94)
podStateDFiltered <- select(podStateD,1,15,16,21)
recceTaskFiltered <- select(recceTask,1,8,272)

```
Define column names
```{r Define column names}
podINS_colNames <- c('timestamp','valid','eastVel','northVel','downVel')
podStateA_colNames <- c('timestamp','mach','gpstime')
podStateD_colNames <- c('timestamp','fov','shutterStatus','sensorID')
recceTask_colNames <- c('timestamp','mpID','groundSpeed')

  
```

```{r Rename}
colnames(podINSFiltered) <- podINS_colNames
colnames(podStateAFiltered) <- podStateA_colNames
colnames(podStateDFiltered) <- podStateD_colNames
colnames(recceTaskFiltered) <- recceTask_colNames
```



#### Valid data  
```{r}
podINSFiltered <- filter(podINSFiltered,valid == "01")
```

#### Combine the Pod INS East / North / Down data
```{r Combine for Meters per Second then convert to Knots}
podINSResult <- mutate(podINSFiltered, combo = eastVel^2 + northVel^2 + downVel^2, ms = sqrt(combo),knots = ms*1.943844)



```

#### Plot the AC Speed  
```{r}
p <- qplot(x = timestamp, y = knots, data = podINSResult,geom = "line")
```

```{r}
summary(podINSResult$knots)
```
```{r}
ggplotly(p)

```








